Discrete Probability for Discrete Quantum Theories

### 2016-05-07 prob.tex
prob.tex Add some comments...

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $\ell$ apart. It is a known fact that
the probability of the needle crossing a line is
$\frac{2}{\pi} = 0.6366197723675814$. We analyze this situation in the
mathematical framework of probability spaces paying special attention
to the resources needed to estimate the probability computationally or
experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles and observe the number $X$ of needles
that cross a line. The sample space can be expressed as the set
$(X|-)^N$ of sequences of characters of length $N$ where each
character is either $X$ to indicate a needle crossing a line or $-$ to
indicate a needle not crossing a line. If $N=3$, the event that
exactly 2 needles cross lines is $\{-XX,X-X,XX-\}$: it has
probability~$\frac{2}{3}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

In an actual experiment with 500 needles, it was found that 321
crossed a line which gives a probability of $0.642$. In a larger
experiment with $3408$ needles, the probability was
$0.6366197193098592$. Comparing the values to the idealized
mathematical value, we see that the observed probability approaches
$\frac{2}{\pi}$ but only if \emph{larger and larger resources} are
expended.

@@

\amr{
From Puri and Ralescu's paper: Fix $n=1$ (only 1-dimension). Conditions on $\mu$:
\begin{itemize}
\item $\mu(E) \neq \emptyset$ for every $E \in \events$, 
\item $1 \in \mu(\Omega)$, 
\item for every disjoint family $A_i$ we have, $\mu(\bigcup_i A_i) =
  \sum_i \mu(A_i)$
\end{itemize}
where the sum of subsets of $\mathbb{R}$ (not $\mathbb{R}^n$ as in the
paper) is as follows. Let $X$ and $Y$ be subsets of $\mathbb{R}$, then
$X + Y = \{ x + y ~|~ x \in X, y \in Y \}$
}

>>>>

Suppose we drop a needle of length $\ell$ onto a floor made of equally
spaced parallel lines a distance $\ell$ apart. It is a known fact that
the probability of the needle crossing a line is
$\frac{2}{\pi} = 0.6366197723675814$. 
\\
\yutsung{Add $\ldots$ at the end? }\\
We analyze this situation in the
mathematical framework of probability spaces paying special attention
to the resources needed to estimate the probability computationally or
experimentally.

To formalize the experiment, we consider an experimental setup
consisting of a collection of $N$ identical needles of length
$\ell$. We throw the $N$ needles and observe the number $X$ of needles
that cross a line. The sample space can be expressed as the set
$(X|-)^N$ \\
\yutsung{Use more physical or mathematical notation?}\\
of sequences of characters of length $N$ where each
character is either $X$ to indicate a needle crossing a line or $-$ to
indicate a needle not crossing a line. If $N=3$, the event that
exactly 2 needles cross lines is $\{-XX,X-X,XX-\}$: it has
probability~$\frac{2}{3}$. Generally, the event that exactly $M$
needles out of the $N$ total needles cross lines is $\frac{M}{N}$.

In an actual experiment with 500 needles, it was found that 321
crossed a line which gives a probability of $0.642$. In a larger
experiment with $3408$ needles, the probability was
$0.6366197193098592$. 
\\
\yutsung{Add $\ldots$ at the end? }\\
Comparing the values to the idealized
mathematical value, we see that the observed probability approaches
$\frac{2}{\pi}$ but only if \emph{larger and larger resources} are
expended.

@@

\amr{
From Puri and Ralescu's paper: Fix $n=1$ (only 1-dimension). Conditions on $\mu$:
\begin{itemize}
\item $\mu(E) \neq \emptyset$ for every $E \in \events$, 
\item $1 \in \mu(\Omega)$, 
\item for every disjoint family $A_i$ we have, $\mu(\bigcup_i A_i) =
  \sum_i \mu(A_i)$
\end{itemize}
where the sum of subsets of $\mathbb{R}$ (not $\mathbb{R}^n$ as in the
paper) is as follows. Let $X$ and $Y$ be subsets of $\mathbb{R}$, then
$X + Y = \{ x + y ~|~ x \in X, y \in Y \}$
}\\
\yutsung{Actually, using $X+Y=\{x+y~|~x\in X,y\in Y\}$ directly
might be more confusing than I thought... Because our example will
be follows:

For $\{0\}=\imposs$ and $(0,\infty)=\poss$, it will become
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & \{0\}\\
\pmeas(\{HH\}) & = & (0,\infty)\\
\pmeas(\{HT\}) & = & \{0\}\\
\pmeas(\{TH\}) & = & (0,\infty)\\
\pmeas(\{TT\}) & = & \{0\}\\
\pmeas(\{HH,HT\}) & = & (0,\infty)\\
\pmeas(\{HH,TH\}) & = & (0,\infty)\\
\pmeas(\{HH,TT\}) & = & (0,\infty)
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & (0,\infty)\\
\pmeas(\{HT,TT\}) & = & \{0\}\\
\pmeas(\{TH,TT\}) & = & (0,\infty)\\
\pmeas(\{HH,HT,TH\}) & = & (0,\infty)\\
\pmeas(\{HH,HT,TT\}) & = & (0,\infty)\\
\pmeas(\{HH,TH,TT\}) & = & (0,\infty)\\
\pmeas(\{HT,TH,TT\}) & = & (0,\infty)\\
\pmeas(\{HH,HT,TH,TT\}) & = & (0,\infty)
\end{array}\end{array}
\]

For $\{0\}=\imposs$ and $(0,1]=\poss$, it will become the following
and gives an extra-value~$(0,2]$.
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & \{0\}\\
\pmeas(\{HH\}) & = & (0,1]\\
\pmeas(\{HT\}) & = & \{0\}\\
\pmeas(\{TH\}) & = & (0,1]\\
\pmeas(\{TT\}) & = & \{0\}\\
\pmeas(\{HH,HT\}) & = & (0,1]\\
\pmeas(\{HH,TH\}) & = & (0,2]\\
\pmeas(\{HH,TT\}) & = & (0,1]
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & (0,1]\\
\pmeas(\{HT,TT\}) & = & \{0\}\\
\pmeas(\{TH,TT\}) & = & (0,1]\\
\pmeas(\{HH,HT,TH\}) & = & (0,2]\\
\pmeas(\{HH,HT,TT\}) & = & (0,1]\\
\pmeas(\{HH,TH,TT\}) & = & (0,2]\\
\pmeas(\{HT,TH,TT\}) & = & (0,1]\\
\pmeas(\{HH,HT,TH,TT\}) & = & (0,2]
\end{array}\end{array}
\]

For $\{0\}=\imposs$, $[0,\frac{1}{2}]=\unlikely$, and $[\frac{1}{2},1]=\likely$,
it will become the following and gives an extra-value~$[\frac{1}{2},\frac{3}{2}]$.
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & \{0\}\\
\pmeas(\{HH\}) & = & [0,\frac{1}{2}]\\
\pmeas(\{HT\}) & = & \{0\}\\
\pmeas(\{TH\}) & = & [\frac{1}{2},1]\\
\pmeas(\{TT\}) & = & \{0\}\\
\pmeas(\{HH,HT\}) & = & [0,\frac{1}{2}]\\
\pmeas(\{HH,TH\}) & = & [\frac{1}{2},\frac{3}{2}]\\
\pmeas(\{HH,TT\}) & = & [0,\frac{1}{2}]
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & [\frac{1}{2},1]\\
\pmeas(\{HT,TT\}) & = & \{0\}\\
\pmeas(\{TH,TT\}) & = & [\frac{1}{2},1]\\
\pmeas(\{HH,HT,TH\}) & = & [\frac{1}{2},\frac{3}{2}]\\
\pmeas(\{HH,HT,TT\}) & = & [0,\frac{1}{2}]\\
\pmeas(\{HH,TH,TT\}) & = & [\frac{1}{2},\frac{3}{2}]\\
\pmeas(\{HT,TH,TT\}) & = & [\frac{1}{2},1]\\
\pmeas(\{HH,HT,TH,TT\}) & = & [\frac{1}{2},\frac{3}{2}]
\end{array}\end{array}
\]

Actually, this approach is more natural to me when I think in measure
theory, but I guess the readers might be confused if write in this
way?}
